The Role of Context in Spatial Region Identification

### 2013-11-15 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

\subsection{Task 1 - Marking Region}
In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.  Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.  There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.  We used two methods to analyze the results inter-annotator agreement and comparing features of CDSRs between stimuli.

Second, we use features of the regions (e.g., area, maximum x-coordinate) to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.

@@

%Kate:  (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with %context,  I changed this SUBSTANTIALLY to something that I think sounds better, but makes a different argument.  Let me know if it is ok.  Really %agreement would indicated shared understanding while variation between contexts would lend support to the importance of context.  Not sure I %captured this
We use two methods to analyze the results of the region drawing task.  First, we use inter-annotator agreement for two reasons: (1) fair to good overall agreement would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).

To measure the agreement between two regions, we use \textit{Cohen's kappa} coefficient, $\kappa$, as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimulus basis. In this calculation, if either of the subjects in a pair entered a no region response for a particular stimulus a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.  Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.

@@

\subsubsection{Predicted Region Changes and Simulation}
\label{sec:simulation}
To assess the claim that changing contexts predictably affects the CDSRs, we identified six expected relationships of region features between stimuli.  For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-x coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).  CDSRs can be dynamic. Thus, in the move to safety scenario, we expect that people will simulate the actions of the agents.  Therefore, we expect the maximum-Y to be larger in the tanks scenario (j) than the turrets (k) because the friendly tanks will pursue the enemy tanks.  Similarly, we expect maximum-Y coordinate to the be same in the two different turrets scenarios (k) and (l), because people would simulate the results of the battle and realize that a single turret will not provide safety against three tanks.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the CDSR features between stimuli.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically affect the perceived extent of a CDSR.  The last two rows of the table evaluate our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.  The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that subjects were simulating actions of other agents in the scene.  In the last row, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).  The average maximum-Y was greater for the turret +1 stimuli indicating that people felt that the single turret provided protection from the enemy tanks.  As discussed in the previous section, we expected some ambiguity in this trial as people likely have different models for the capabilities of the different units in this scenario.

@@

To assess the claim that changing contexts predictably affects the CDSRs, we can compare the ratings of points between stimuli.  In the classroom domain, we assume that part of the definition of the front of the room is where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the U-shaped scenario (c) than the rows scenarios (e) and (f).  This is confirmed in the results from Table \ref{tab:point-likert-scores}.  In the military domain, in stimulus (i), the friendly tanks are positioned between the P2 point and the enemy tanks, whereas, in stimulus (j), the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.  Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). This is confirmed in Table \ref{tab:point-likert-scores} and is statistically significant $(p<.05)$.  To assess the extent of participants use of simulation in determining the context, we compare the subject's ratings of the \textit{no-man's land} point between the tanks (j) and the turrets (k) stimuli as well as between the turrent (k) and turrets +1 (l).  The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimuli (j) than (k).  The single turret should offer only temporary protection against the enemy tank force, therefore it should not increase the safety rating.  Neither of these results were found in Table \ref{tab:point-likert-scores}.  As discussed in Section \ref{sec:simulation}, we expected ambiguity with the turrets +1 stimuli.  The difference in the rating of P2 between the tanks and turrets scenario provides two alternative explanations regarding how participants interpreted the context:  (1) the turrets were viewed as superior defensive units and, thus, better able to provide safety, or (2) the turrets being fixed could not move away, and thus, establish a more permanent CDSR for safety.  Further experiments are needed to explore these hypotheses.

@@

Regions are central to qualiative spatial reasoning (QSR)~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points seem unlikely to model the regions with sufficient flexibility.

>>>>

\subsection{Task 1 - Marking Region}
In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.  Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.  There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.  We analyzed the results in two ways: measuring inter-annotator agreement  and comparing features of CDSRs between stimuli.

@@

%Kate:  (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with %context,  I changed this SUBSTANTIALLY to something that I think sounds better, but makes a different argument.  Let me know if it is ok.  Really %agreement would indicated shared understanding while variation between contexts would lend support to the importance of context.  Not sure I %captured this
We tested two hypotheses concerning inter-annotator agreement. First, we expect fair to good overall agreement indicating that subjects share the interpretation of the CDSRs in the given contexts.  Second, we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).

To measure the inter-annotator agreement between two subjects, we use \textit{Cohen's kappa} coefficient, $\kappa$, as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimulus basis. In this calculation, if either of the subjects in a pair entered a no region response for a particular stimulus a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.  Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.

@@

\subsubsection{Predicted Region Changes Across Stimuli}
\label{sec:simulation}
To assess the claim that changing contexts predictably affects the CDSRs, we identified six expected relationships of region features between stimuli.  For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-X coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).  CDSRs can be dynamic. Thus, in the move to safety scenario, we expect that people will simulate the actions of the agents.  Therefore, we expect the maximum-Y to be larger in the tanks scenario (j) than the turrets (k) because the friendly tanks will pursue the enemy tanks.  Similarly, we expect maximum-Y coordinate to the be same in the two different turrets scenarios (k) and (l), because people would simulate the results of the battle and realize that a single turret will not provide safety against three tanks.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the CDSR features between stimuli.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically affect the perceived extent of a CDSR.  The last two rows of the table evaluate our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.  The comparison between the maximum-Y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that subjects were simulating actions of other agents in the scene.  In the last row, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).  The average maximum-Y was greater for the turret +1 stimuli indicating that people felt that the single turret provided protection from the enemy tanks.  As discussed in the previous section, we expected some ambiguity in this trial as people likely have different models for the capabilities of the different units in this scenario.

@@

To assess the claim that changing contexts predictably affects the CDSRs, we can compare the ratings of points between stimuli.  In the classroom domain, we assume that part of the definition of the front of the room is where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the U-shaped scenario (c) than the rows scenarios (e) and (f).  This is confirmed in the results from Table \ref{tab:point-likert-scores}.  In the military domain, in stimulus (i), the friendly tanks are positioned between the P2 point and the enemy tanks, whereas, in stimulus (j), the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.  Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). This is confirmed in Table \ref{tab:point-likert-scores} and is statistically significant $(p<.05)$.  To assess the extent of participants use of simulation in determining the context, we compare the subject's ratings of the \textit{no-man's land} point between the tanks (j) and the turrets (k) stimuli as well as between the turrent (k) and turrets +1 (l).  The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimulus (j) than stimulus (k).  The single turret should offer only temporary protection against the enemy tank force, therefore it should not increase the safety rating.  Neither of these results were found in Table \ref{tab:point-likert-scores}.  As discussed in Section \ref{sec:simulation}, we expected ambiguity with the turrets +1 stimuli.  The difference in the rating of P2 between the tanks and turrets scenario provides two alternative explanations regarding how participants interpreted the context:  (1) the turrets were viewed as superior defensive units and, thus, better able to provide safety, or (2) the turrets being fixed could not move away, and thus, establish a more permanent CDSR for safety.  Further experiments are needed to explore these hypotheses.

@@

Regions are central to qualiative spatial reasoning (QSR)~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai,kelleher/costello:09} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points seem unlikely to model the regions with sufficient flexibility.
